Actuator line modeling of vertical-axis turbines

### 2016-05-02 paper.tex
Mention NS based vortex method
Add clarification to BE diagram

The ALM, originally developed by Sorensen and Shen \cite{Sorensen2002}, has
become popular for modeling axial-flow or horizontal-axis turbines, and has been
shown in blind tests to be competitive with blade-resolved
CFD~\cite{Krogstad2013, Pierella2014}. The ALM combined with large eddy
simulation has become the state-of-the-art for modeling entire wind
farms~\cite{Archer2013, Churchfield2012, Sorensen2015, Fleming2013,
Fleming2014}. Like other blade element techniques, the effectiveness of the ALM
for AFTs is in part due to the quasi-steady nature of the flow in the blade
reference frame, and the relatively rare occurrence of stall.

@@

    \caption{Vector diagram of velocity and forcing on a cross-flow turbine
    blade element. Note that the free stream flow is from top to bottom, the
    blade chord is in line with the tangential velocity (i.e., zero preset
    pitch), and the drag vector is magnified by a factor of two (approximately,
    relative to the lift vector) to enhance visibility.}

>>>>

The ALM, originally developed by Sorensen and Shen \cite{Sorensen2002}, has
become popular for modeling axial-flow or horizontal-axis turbines, and has been
shown in blind tests to be competitive with blade-resolved
CFD~\cite{Krogstad2013, Pierella2014}. The ALM combined with large eddy
simulation has become the state-of-the-art for modeling entire wind
farms~\cite{Archer2013, Churchfield2012, Sorensen2015, Fleming2013,
Fleming2014}. Like other blade element techniques, the effectiveness of the ALM
for AFTs is in part due to the quasi-steady nature of the flow in the blade
reference frame, and the relatively rare occurrence of stall. Note that a
similar method can be used with the Navier--Stokes equations in
vorticity--velocity form~\cite{Scheurich2011b}, which may be more efficient when
vorticity is confined to a relatively small fraction of the domain, e.g., for a
standalone turbine with a uniform laminar inflow.

@@

    \caption{Vector diagram of velocity and forcing on a cross-flow turbine
    blade element. Note that the free stream velocity $U_\infty$ is from top to
    bottom (identical to $U_in$ for purely geometric calculations), the blade
    chord is in line with the tangential velocity (i.e., zero preset pitch,
    which would offset the geometric angle of attack $\alpha$), and the drag
    vector is magnified by a factor of two (approximately, relative to the lift
    vector) to enhance visibility.}
